Verification of the CVM algorithm with a Functional Probabilistic Invariant

### 2025-02-26 paper/sec/intro.tex
add more line count to show benefit of invariant

We have also mechanized the transformation-based CVM proof by Chakraborty et al.~\cite{chakraborty2022,chakraborty2023}.
As a rough point of comparison, verification of the CVM algorithm using our new technique requires only \locnew~lines in Isabelle/HOL~\cite{nipkow2002}, while the original proof required \locold~lines.

The rest of this paper is organized as follows.
\cref{sec:background} provides background information on randomized algorithms, in particular on their semantics in Isabelle/HOL.
\cref{sec:invariants} introduces our new technique and explains how probabilistic loop invariants can be used to establish tail bounds for the original CVM algorithm.
\cref{sec:negdep} introduces the concept of negative association and our new total and unbiased variant of the CVM algorithm.
\cref{sec:formalization} presents the formalization of both variants of the algorithm, and \cref{sec:formalization_neg_dep} describes our new formalized library on negatively associated random variables.
\cref{sec:transformation_based_proof} discusses some challenges faced in our alternative verification of CVM using the transformation-based proof by Chakraborty et al.
The final sections conclude with an overview of related work and a summary of our results.

>>>>

Our mechanization is carried out in Isabelle/HOL~\cite{nipkow2002}, where we generically formalized the invariant-based argument underlying both CVM variants in 753 lines.
Specializing to the original CVM algorithm took an additional 250 lines, while the new variant took 191 more lines (with 2974 lines for the supporting negative association library).
We also mechanized the transformation-based proof by Chakraborty et al.~\cite{chakraborty2022,chakraborty2023}.
As a rough point of comparison showing the simplicity of our new proof, verification of the CVM algorithm using our new technique required only \locnew~(753 + 250)~lines, while the original proof required \locold~lines.

The rest of this paper is organized as follows.
\cref{sec:background} provides background information on randomized algorithms, in particular on their semantics in Isabelle/HOL.
\cref{sec:invariants} introduces our new technique and explains how probabilistic loop invariants can be used to establish tail bounds for the original CVM algorithm.
\cref{sec:negdep} introduces the concept of negative association and our new total and unbiased variant of the CVM algorithm.
\cref{sec:formalization} presents the formalization of both variants of the algorithm, and \cref{sec:formalization_neg_dep} describes our new formalized library on negatively associated random variables.
\cref{sec:transformation_based_proof} discusses some challenges faced in our alternative verification of CVM using the transformation-based proof by Chakraborty et al.
The final sections present related work and a summary of our results.
